% !TEX root = root.tex
\documentclass[11pt, a4paper]{article} % Removed twocolumn
\usepackage[colorlinks=true, urlcolor=blue, linkcolor=red]{hyperref}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage[margin=1in]{geometry}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning, calc}

\title{Predicting the likelihood of a TikTok video going viral using intrinsic and extrinsic signals}
\author{Oscar Rodriguez, Mahda Soltani, Ela Naz Sigin}
\date{March 10, 2026}

\begin{document}

\maketitle

\begin{abstract}
    Social media platforms, like TikTok have become one of the dominant
    channels through which information, entertainment, and political and
    cultural trends propagate at scale. A defining feature of these
    platforms is the extreme skewness in attention: while the vast majority
    of uploaded videos receive little to no exposure, a small fraction
    achieve \textit{viral} reach, accumulating orders of magnitude more
    views within a short time window. This heavy-tailed distribution of
    engagement naturally raises the question of how to operationalize and
    predict “virality” in a principled way.

    This investigation explores the use of Machine Learning techniques to
    pin down a formal framework for the task of predicting the likelihood of
    a TikTok video going viral using both intrinsic (e.g. video content) and
    extrinsic features (e.g. video author).
\end{abstract}

\section*{Introduction}
\dots

\section*{Methods}

This section goes over the different aspects of the Machine Learning pipeline
that we crafted for the given task, starting from data processing all the way
to the model architecture and reasoning behind each of the decisions.

\subsection*{Data collection and processing}

We started off with a base dataset called
\href{https://huggingface.co/datasets/The-data-company/TikTok-10M}{TikTok-10M}
from which we sourced a rich subset of examples that we complemented with a
scraping step to download the videos and publisher accounts to extract
additional signals (e.g. video bytes, aspect ratio, etc.) as well
as account information (e.g. account statistics, publishing history, etc.).

With all this raw data at hand, we produced a
\href{https://huggingface.co/datasets/rodmosc/viral}{curated dataset} for
our task. The following tables contain a detailed documentation of dataset
dimensions, encompassing intrinsic content signals and extrinsic author/metadata
signals we deemed relevant to the problem.

\begin{table}
\centering
\small
\begin{tabularx}{\textwidth}{|l l X l|}
    \hline
    \textbf{Dimension} & \textbf{Type} & \textbf{Description} & \textbf{Extraction / Treatment} \\
    \hline
    \multicolumn{4}{|l|}{\textbf{User / author signals}} \\
    \hline
    user\_id/username & ID/String & Unique identifier and handle of the creator & Scraped (TikTok web profile) \\
    user\_language & String & Configuration language of the account & Base dataset \\
    city & String & City where the video was posted from & Base dataset \\
    user\_verified & Boolean & Verification status of the author & Scraped (TikTok web profile) \\
    is\_private & Boolean & Whether the account is set to private & Scraped (TikTok web profile) \\
    author\_stats & Float & Follower, heart, video, and friend counts & Scraped; $\log(1+x)$ scaled \\
    account\_create\_time & Timestamp & When the user account was first created & Scraped (unix epoch) \\
    \hline
\end{tabularx}
\end{table}

\begin{table}
\centering
\small
\begin{tabularx}{\textwidth}{|l l X l|}
    \hline
    \multicolumn{4}{|l|}{\textbf{Video content \& metadata}} \\
    \hline
    description & String & User-generated caption and hashtags & Base dataset \\
    create\_time & Timestamp & Exact time of video publication & Base dataset (UTC) \\
    temporal\_features & Float & Hour of day and Day of week & Sine/cosine cyclic encoding \\
    is\_ad & Boolean & Flag for paid/sponsored content & Base dataset \\
    duration & Integer & Total length of the video in seconds & Base dataset \\
    challenges & String & Associated hashtag challenges or trends & Base dataset \\
    video\_bytes & Binary & Scraped video resized to 256x256 @ 1fps & \texttt{yt-dlp} + \texttt{moviepy} \\
    detected\_objects & String & Objects detected in video & pretrained facebook/detr-resnet-50 \\
    \hline
\end{tabularx}
\end{table}

\begin{table}
\centering
\small
\begin{tabularx}{\textwidth}{|l l X l|}
    \hline
    \multicolumn{4}{|l|}{\textbf{Audio \& music specs}} \\
    \hline
    music\_info & String & Title, album, and author name of the audio & \texttt{yt-dlp} payload metadata \\
    music\_play\_url & String & URL source for the audio file & Base dataset \\
    \hline
\end{tabularx}
\end{table}

\begin{table}
\centering
\small
\begin{tabularx}{\textwidth}{|l l X l|}
    \hline
    \multicolumn{4}{|l|}{\textbf{Technical / video specs}} \\
    \hline
    resolution & Integer & Video width and height in pixels & \texttt{yt-dlp} (Best format) \\
    aspect\_ratio & Float32 & Ratio of width to height & Calculated from payload \\
    filesize & Integer & Total storage size of the video file & \texttt{yt-dlp} payload metadata \\
    codecs & String & Video (vcodec) and audio (acodec) formats & \texttt{yt-dlp} payload metadata \\
    vq\_score & Float64 & Visual quality score & Base dataset \\
    \hline
\end{tabularx}
\end{table}

\begin{table}
\centering
\small
\begin{tabularx}{\textwidth}{|l l X l|}
    \hline
    \multicolumn{4}{|l|}{\textbf{Engagement stats \& target labels}} \\
    \hline
    interaction\_counts & Integer & Plays, likes, comments, shares, and saves & Scraped + base dataset \\
    engagement\_score & Float64 & Weighted sum of interactions per view & Engineered; Weighted calculation \\
    view\_velocity & Float64 & Views accumulated per hour since creation & Engineered; $\log(1+x)$ scaled \\
    is\_viral & Binary & Classification ($P=0.95$) of engagement & Engineered (Quantile threshold) \\
    \hline
\end{tabularx}
\end{table}

\subsection*{Task definition}

The definition of virality (is\_viral) we produced is the target dimension we
optimize for. It is meant to capture outliers relative to any given account
stack-ranked in absolute terms with the goal of producing a signal that's useful
for any account size.

In order to recreate a mixed signal that fits the previous definition, we
compute two scores for each video, \textbf{Engagement} and \textbf{View
Velocity}.

\begin{align}
    \text{Engagement}_i &= \frac{\text{share\_count} \times w_0 + \text{save\_count} \times w_1 + \text{comment\_count} \times w_2 + \text{like\_count} \times w_3}{\text{play\_count}} \\
\end{align}

Where $Engagement$ is meant to represent a rate of amount of "interactions" a
video had relative to its play count, making the score relative to the account
size. As shown in the following graph, its expected to assume that accounts
with a larger audience will most likely outperform its counterparts, making
these videos seem more "viral" in absolute measurements.

% TODO: Insert graph showing distribution of play count binned by account size.

What this signal is meant to capture instead is a degree of "engagement" that's
normalized to the audience size of any given account. To convey the idea, a
small account with a few hundred users that produces a video that generates
thousands of interactions would have a comparable engagement to an acconut with
millions of users that accumulated hundredths of thousands of interactions
instead.

\begin{align}
    \text{View Velocity}_i &= \frac{\text{play\_count}}{\text{delta\_hours}}
\end{align}

Where \textbf{View Velocity} normalizes the value of $play\_count$ by the total
amount of time a video has been live collecting interactions. A video with a
high "view velocity" is collecting interactions "faster" relative to the rest.

% TODO: Insert graph showing distribution of play count binned by delta hours.

We add together both measurements to produce a signal $S_i$ for each video that
we use to stack-rank them in absolute $S_i$ terms and consider the top $5\%$ as
"viral":

\begin{align}
    S_i &= \frac{\text{Engagement}_i}{\text{Engagement}_{max}} + \frac{\text{View\_Velocity}_i}{\text{View\_Velocity}_{max}} \\
    y_{\text{is\_viral}}^{(i)} &= 
    \begin{cases} 
      1 & \text{if } S_i \geq F_S^{-1} (0.95) \\
      0 & \text{Otherwise}
    \end{cases}
\end{align}

The reason we approached the concept of virality as described is rooted in
convenience and usefulness. There are just too many dynamic forces that could
influence the level of virality a video could reach. Many things like the
current social discourse, ongoing political and economic events, or just sheer
luck (being in the right place at the right time) could determine the success of
any given video.

Trying to capture such a broad range of "reasons" is likely near impossibility
limits either due to data or resource constraints.

Instead we try to condense a signal that's useful even for small accounts with
the available data. We're thinking about small businesses/ content creators
looking to amplify their social presence. Rolling out a video that's a couple of
folds more viral relative to their expected baseline is considered a win!

\subsection*{Model architecture}

The architecture implements a mixture of state-of-the-art ML processing
techniques for the different data modalities fed into the model; all these get
fed through a late fusion module to finally produce 2 regression heads for the
engagement and view velocity scores and 1 classification head for the virality
class. We ultimately stitch them together to compute a final weighted loss to
have the different output targets contribute in different quantities.

\begin{figure}[ht]
\centering
\resizebox{\textwidth}{!}{
\begin{tikzpicture}[
    node distance=1.2cm and 1.5cm,
    base_style/.style={rectangle, draw, rounded corners, thick, align=center, inner sep=10pt},
    block/.style={base_style, fill=white, text width=3.2cm, minimum height=1.2cm},
    input/.style={base_style, fill=gray!10, text width=3.2cm, minimum height=1.6cm},
    branch/.style={base_style, fill=blue!5, text width=3.4cm, minimum height=1.4cm},
    head/.style={base_style, fill=orange!5, text width=3.4cm, minimum height=1.4cm},
    loss/.style={base_style, fill=red!5, text width=4.2cm, minimum height=1.8cm},
    arrow/.style={-Stealth, thick}
]
% --- Inputs ---
\node[input] (text_in) {\textbf{Text input}\\ \small \texttt{input\_ids} \\ \small (batch, 256)};
\node[input, below=of text_in] (video_in) {\textbf{Video input}\\ \small \texttt{pixel\_values} \\ \small (batch, 16, 3, 224, 224)};
\node[input, below=of video_in] (tab_in) {\textbf{Tabular input}\\ \small \texttt{tabular\_features} \\ \small (batch, 19)};
% --- Modules ---
\node[branch, right=of text_in] (text_mod) {\textbf{Text model}\\ \small Frozen DistilBERT \\ \small (hidden: 768)};
\node[branch, right=of video_in] (video_mod) {\textbf{Video model}\\ \small Frozen VideoMAE \\ \small (hidden: 768)};
\node[branch, right=of tab_in] (tab_mod) {\textbf{Tabular model}\\ \small 2-layer MLP \\ \small (d\_model: 512)};
% --- Fusion ---
\node[block, right=1.2cm of video_mod, fill=green!5, text width=3.8cm] (fusion) {
    \textbf{Late fusion}\\ \small (dim=2048) \\
};
% --- Heads ---
\node[head, right=1.5cm of fusion] (class_head) {\textbf{Classification head}\\ \small Virality logit \\ \small (batch, 1)};
\node[head, above=0.5cm of class_head] (eng_head) {\textbf{Engagement head}\\ \small Regression logit \\ \small (batch, 1)};
\node[head, below=0.5cm of class_head] (vel_head) {\textbf{Velocity head}\\ \small Regression logit \\ \small (batch, 1)};
% --- Loss aggregator ---
\node[loss, right=1.8cm of class_head] (total_loss) {
    \textbf{Multi-task loss}\\
    \scriptsize $\mathcal{L}_{\text{total}} = \lambda_{\text{reg}}\mathcal{L}_{\text{Huber}} + \lambda_{\text{cls}}\mathcal{L}_{\text{BCE}}$ \\
    \small $\lambda_{\text{reg}}=0.1, \lambda_{\text{cls}}=0.9$
};
% --- Arrows ---
\draw[arrow] (text_in) -- (text_mod);
\draw[arrow] (video_in) -- (video_mod);
\draw[arrow] (tab_in) -- (tab_mod);
\draw[arrow] (text_mod.east) -- ++(0.5,0) |- (fusion.west);
\draw[arrow] (video_mod.east) -- (fusion.west);
\draw[arrow] (tab_mod.east) -- ++(0.5,0) |- (fusion.west);
\draw[arrow] (fusion.east) -- ++(0.5,0) |- (eng_head.west);
\draw[arrow] (fusion.east) -- (class_head.west);
\draw[arrow] (fusion.east) -- ++(0.5,0) |- (vel_head.west);
% --- Loss connections ---
\draw[arrow, dashed, gray] (eng_head.east) -- ++(0.4,0) |- (total_loss.160);
\draw[arrow, dashed, gray] (vel_head.east) -- ++(0.4,0) |- (total_loss.200);
\draw[arrow, dashed, gray] (class_head.east) -- (total_loss.west);
\end{tikzpicture}
}
\caption{End-to-end architecture: features are extracted from multimodal inputs,
fused into a shared latent space, and passed to specialized heads. The final
objective is a weighted combination of Huber regression loss and weighted binary
cross-entropy (BCE) loss.}
\end{figure}

\subsubsection*{Data modalities}

\subsubsection*{Text input}

The text input is a concatenation of the different text dimensions in the
following form:

\begin{quote}
    "[CLS] [TITLE] ... [MUSIC] ... [CHALLENGES] ... [OBJECTS] ... [LOC] ..."
\end{quote}

That we populate with the example's data and feed this into a pretrained, frozen
\href{https://huggingface.co/distilbert/distilbert-base-uncased}{DistilBERT}
backbone to extract a 768-dimensional semantic representation from the [CLS]
token, capturing the semantics of all textual components from the example in a
compact representation.

\subsubsection*{Video input}

During the dataset composition step, the original videos were downsized and
downsampled to 224x224 pixels at 1 FPS to reduce processing requirements. From
that transformed version, we sample 16 frames that we feed into a pretrained,
frozen \href{https://huggingface.co/MCG-NJU/videomae-base}{VideoMAE} model to
encode the video into a spatial-temporal embedding space.

We then apply global average pooling across the output tokens to yield a
768-dimensional embedding that captures both visual features and temporal
dynamics.

\subsubsection*{Tabular input}

The tabular branch processes 19 heterogeneous features including author signals,
temporal metadata, and video quality scores. To handle the varying scales of
these inputs, we apply $log(1+x)$ scaling to high-variance author statistics and
sine/cosine cyclic encoding to temporal features like hour of day to preserve
periodic relationships. These are fed into a 2-layer MLP with layer
normalization and dropout, projecting them into a learnable 512-dimensional
space.

\subsubsection*{Late fusion}

In the late fusion module, we concatenate the 768-dimension text embedding,
the 768-dimension video embedding, and the 512-dimension tabular output into a
single 2048-dimensional vector. This combined representation is passed through a
shared fusion MLP, which maps the multimodal features into a unified latent
space of $d\_model=512$. This shared representation allows the model to learn
complex inter-modal correlations before branching into specialized tasks.

\subsubsection*{Multi-head targets}

We employ a multi-head architecture to simultaneously model continuous and
categorical virality signals. Each head progressively halves the upstream
$d\_model$ to reduce dimensionality while increasing specialization:

\begin{itemize}
    \item Engagement and velocity heads: Two regression heads predict the
    continuous engagement and view velocity scores.
    \item Classification head: A single head predicts the probability of
    virality, defined by the 95th percentile of the combined engagement and
    velocity distribution.
\end{itemize}

\subsubsection*{Weighted loss target}

The loss target is a composite of the different upstream heads, weighted to
prioritize the primary classification task while using the regression heads as
auxiliary signals:

\begin{itemize}
    \item Loss composition: We use Huber loss for the regression targets to
    remain robust against the extreme outliers common in viral distributions,
    and weighted BCE for the classification target.
    \item Weighting strategy: We set $\lambda_{cls}=0.7$ and
    $\lambda_{reg}=0.3$. This heavy weighting on classification ensures the
    model focuses on detecting the rare viral events (upweighted by a factor of
    $virality\_threshold / (1 - virality\_threshold)$ in the BCE loss), while
    the regression task provides a continuous "popularity" signal that
    stabilizes the shared latent space.
\end{itemize}

\section*{Experiments and results}

\dots

\section*{Next steps}

\dots

\section*{References}

\dots

\end{document}